\documentclass{amsart}

\usepackage{xcolor}

\setlength{\parindent}{0pt}
\setlength{\parskip}{\baselineskip}

\raggedbottom

\begin{document}
\begin{center}
Science Atlantic Math Competition 2021 Solutions
\end{center}

\textbf{Problem 1}\\
True or false? A $5$-$5$-$6$ triangle has the same area as a $5$-$5$-$8$ triangle.

\textit{Solution}

\textbf{Problem 2}\\
True or false? $31^{11} < 17^{14}$

\textit{Solution}

\textbf{Problem 3}\\
True or false? $\log_10{550} < 2.5$.

\textit{Solution}

\textbf{Problem 4}\\
True or false? There are exactly $10!$ seconds in $6$ weeks.

\textit{Solution}

\textbf{Problem 5}\\
True or false? There is a permutation of $\{0,\dots,9\}$ such that the sum of any three consecutive numbers is $<15$.

\textit{Solution}

\textbf{Problem 6}\\
True or false? The area in the first quadrant enclosed by the graphs of $y = x^{2023}$ and $x^{1/2023}$ is $>0.999$.

\textit{Solution}

\textbf{Problem 7}\\
Alice and Bob play a game in which Alice tosses $2023$ fair coins and Bob tosses $2022$ fair coins.
If Alice obtains a greater number of heads, she wins the game; otherwise Bob wins.
True or false? Bob has a greater probability of winning the game than does Alice.

\textit{Solution}

\textbf{Problem 8}\\
A farmer owns a triangular field, as shown below ({\color{red}to be added later}).
The farmer figures that $5$ sheep can graze in the west field, $10$ in the south field, and $8$ in the east field.
Assume that all sheep eat the same amount of grass.
True or false? At least $21$ sheep can graze in the north field.

\textit{Solution}

\textbf{Problem 9}\\
Suppose that
\[ \tan(x) + \cot(x) + \sec(x) + \csc(x) = 2023. \]
True or false? $\sin(x) + \cos(x) \geq \frac{2}{\sqrt{3}}$.

\textit{Solution}

\textbf{Problem 10}\\
True or false? When a prime number is divided by $60$, the remainder is either $1$ or a prime number.

\textit{Solution}\\
$109$ is prime, but the remainder upon division by $60$ is $49 = 7^2$.

\textbf{Problem 11}\\
Two people plan to meet at a given place between noon and 1PM under the agreement that neither person will wait longer than $12$ minutes for the other.
Assume that each person chooses their time of arrival at random.
True or false? The probability that a meeting takes place is $\geq 1/3$.

\textit{Solution}

\textbf{Problem 12}\\
A road crew can build a section of highway in $3$ days with their present supply of machines.
With $3$ more machines, however, they could do the job in only $2$ days.
True or false? If the road crew had only one machine, it would take more than two weeks to build the highway?

\textit{Solution}\\
The number of machines is inversely proportional to time.
Let $m$ be the current number of machines; then $m + 3 = 3m/2$ and $m = 6$.
If the crew had only one machine, it would take $3 \cdot 6 = 18$ days, which is more than two weeks.

\textbf{Problem 13}\\
The numbers $1,\dots,6$ are assigned to the six faces of a cube, one number to each face.
For each of the eight vertices of the cube, compute the product of the numbers assigned to the three faces incident with that vertex, and then let $S$ be the sum of the eight products.
True or false? Regardless of how the numbers are assigned to the faces, $S < 321$.

\textit{Solution}

\textbf{Problem 14}\\
Let $A$ and $B$ be $2 \times 2$ matrices with real entries.
True or false? If $AB = 0$ then $BA = 0$.

\textit{Solution}\\
\[ \begin{bmatrix} 0 & 1 \\ 0 & 0 \end{bmatrix} \begin{bmatrix} 1 & 0 \\ 0 & 0 \end{bmatrix} = \begin{bmatrix} 0 & 0 \\ 0 & 0 \end{bmatrix} \]
but
\[ \begin{bmatrix} 1 & 0 \\ 0 & 0 \end{bmatrix} \begin{bmatrix} 0 & 1 \\ 0 & 0 \end{bmatrix} = \begin{bmatrix} 0 & 1 \\ 0 & 0 \end{bmatrix} \]

\textbf{Problem 15}\\
True or false? There are at least $280$ positive integer divisors of $20^{23}$ which are perfect squares.

\textit{Solution}

\textbf{Problem 16}\\
True or false? It is possible to partition the set $\{1,\dots,15\}$ into $5$ disjoint $3$-element subsets $S_k = \{a_k,b_k,c_k\}$ such that $a_k + b_k = c_k$ for all $k \in \{1,\dots,5\}$.

\textit{Solution}

\textbf{Problem 17}\\
Let $f(x) = \ln(\sqrt{1 + e^{\cos(x)}})$.
True or false? $f(0)f'(0)f''(0) = -1$.

\textit{Solution}

\textbf{Problem 18}\\
True or false? The value of $\sqrt{9 + \sqrt{9 + \sqrt{9 + \sqrt{9 + \dots}}}}$ is less than $3.5$.

\textit{Solution}

\textbf{Problem 19}\\
A long round wire has a cross-section with a radius of $1$ millimetre.
The wire is tightly wound into a spherical shape, resulting in a ball of radius $20$ centimetres.
True or false? When unravelled, the wire will span (at least) the entire length of the Confederation Bridge (12.9 kilometres).

\textit{Solution}

\textbf{Problem 20}\\
The golden ratio $\phi = (1 + \sqrt{5})/2$ has the remarkable property that its reciprocal is equal to its fractional part.
True or false? The golden ratio is the only positive real number with this property.

\textit{Solution}

\textbf{Problem 21}\\
A finite set $S$ of real numbers has the following properties:
\begin{itemize}
\item $1 \not\in S$ and $2001 \not\in S$,
\item The mean of $S \cup \{1\}$ is $13$ less than the mean of $S$, and
\item The mean of $S \cup \{2001\}$ is $27$ more than the mean of $S$.
\end{itemize}
True or false? There are at least $50$ numbers in $S$.

\textit{Solution}

\textbf{Problem 22}\\
True or false? The equation $x^3 = 3x^2 + 1$ has exactly three real solutions.

\textit{Solution}

\textbf{Problem 23}\\
True or false?
\[ \frac{1}{2} \times \frac{3}{4} \times \dots \times \frac{99}{100} < \frac{1}{10}. \]

\textit{Solution}

\textbf{Problem 24}\\
Fifteen points $P_1,\dots,P_{15}$ are equally spaced around a circle.
True or false?
The size of the angle $P_1 P_3 P_7$ is less than $107$ degrees.

\textit{Solution}

\textbf{Problem 25}\\
Let $f : \mathbb{R} \to \mathbb{R}$ be a continuous function having the property that $f(x) \cdot f(f(x)) = 1$ for all real numbers $x$.
Moreover, suppose that $f(4046) = 4045$.
True or false? $f(2023) < 2023$.

\textit{Solution}

\textbf{Problem 26}\\
Suppose that $(1 + x + x^2)^{2023} = a_0 + \dots + a_{4046}x^{4046}$.
True or false? The integer $a_1 + a_3 + a_5 + \dots + a_{4045}$ is odd.

\textit{Solution}

\textbf{Problem 27}\\
The integer $99$ was multiplied by an integer to obtain the $7$-digit integer $62ab427$, but the digits $a,b$ were illegible.
True or false? The only possible values for the illegible digits are $a = 2$ and $b = 4$.

\textit{Solution}

\textbf{Problem 28}\\
True or false?
\[ \int_0^{\sqrt{\pi/6}} \sin(x^2) \ dx + \int_{-\sqrt{\pi/6}}^{\sqrt{\pi/6}} x^2 \cos(x^2) \ dx < \frac{1}{2\sqrt{2}}. \]

\textit{Solution}

\textbf{Problem 29}\\
Alice drives a truck that gets $10$ kilometres per litre.
Bob drives a small car that gets $40$ kilometres per litre.
Each week, Alice and Bob drive the same distance.
Alice decides to trade her truck for a new one that gets $15$ kilometres per litre.
Bob now wants to trade his car for one which gets $x$ kilometres per litre so that his fuel savings over a week will match those of Alice.
True or false? $x \leq 80$.

\textit{Solution}

\textbf{Problem 30}\\
Let the function $f : \mathbb{R} \to \mathbb{R}$ be defined as follows: $f(x) = \min\{2x + 2,-\frac{3}{4}x + 7\}$.
True or false? The maximum value of $f(x)$ is at least $5$.

\textit{Solution}

\end{document}